---
title: "Riparian Classification"
author: "Hailey Eckstrand, FAIB"
date: "2024-06-19"
output: 
  html_document:
    css: custom.css
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Riparian Classification

This document outlines data sources, methods and assumptions of how the riparian classification dataset was modeled and width of buffers for said classes. Analysis was performed in June, 2024 by FAIB.

Contact: Hailey.Eckstrand@gov.bc.ca  

The riparian classes were derived using the following input datasets and methodology. 

### Data Sources

Data accessed June, 2024

BCGW Layers:

* FWA Wetlands, metadata [link](https://catalogue.data.gov.bc.ca/dataset/freshwater-atlas-wetlands)
* FWA Lakes, metadata [link](https://catalogue.data.gov.bc.ca/dataset/freshwater-atlas-lakes)
* FWA Stream Network, metadata [link](https://catalogue.data.gov.bc.ca/dataset/freshwater-atlas-stream-network)
* FWA Rivers, metadata [link](https://catalogue.data.gov.bc.ca/dataset/freshwater-atlas-rivers)
+ Community Watersheds - Current, metadata [link](https://catalogue.data.gov.bc.ca/dataset/community-watersheds-current)
+ BEC map, metadata [link](https://catalogue.data.gov.bc.ca/dataset/bec-map)
+ Mapped Floodplains in BC (Historical), metadata [link](https://catalogue.data.gov.bc.ca/dataset/mapped-floodplains-in-bc-historical-study-area-limits)

Externally provided data sources:

+ Fish Habitat Accessibility MODEL to determine fish presence/absence (obtained from: Simon Norris via Craig Mount)
+ Modeled channel width (obtained from: Simon Norris via Craig Mount)
+ FREP riparian 2006 - 2023 monitoring data

<details>
  <summary> Expand for further details about data sources.  </summary>
  <br>
* __FREP riparian 2006 - 2023 monitoring data__ - Lisa Nordin has provided an abbreviated version of the data. Sensitive fields were removed and it was recommended to use Albers for the spatial coordinates as the data specialist ensures Albers is correct.
  + Filename: `data/input/June13_Riparian data_for_FAIB.xlsx`
  
* __Fish Habitat Accessibility MODEL__ - Fish presence/absence is required to determine stream classification. The provincial BC Fish Passage database was obtained with the approval of Craig Mount. Approval was granted provided the usage is not for operational purposes as the model is only appropriate for use as a broad landscape analysis tool. The dataset was provided by Simon Norris, the Fish Passage contractor. The dataset is identical to the FWA stream network table but divided into smaller segments. The dataset contains linear_feature_id, allowing it to be related to the FWA stream network table.
  + Filename: `data/input/fishpassage.gpkg.zip`

* __Channel width__ - Also provided by Craig Mount. The channel width model is invalid when the contributing area is outside BC. There is no field that identifies when contributing area extends to outside BC. Note from Simon, June 19, 2024 - "Some progress is being made on that though - turns out Daniel W at the Cultus Lake lab has done the required trans-boundary stream work. He is sharing the data - I haven't seen it yet but I hope to be able to incorporate fixed per-stream upstream area numbers in fwapg sometime soon." The dataset was also provided by Simon Norris. Dataset contains linear_feature_id, allowing them to be related to the FWA stream network table. For this analysis, stream reaches with a contributing area outside BC were identified manually and set to NULL. 
  + Filename: `data/input/fwa_stream_networks_channel_width.csv.gz`


Description of how channel width was calculated:  
•	_“To obtain estimates of channel width upstream of crossing locations, where available, bcfishpass was utilized to pull average channel gradients from Fisheries Information Summary System (FISS) site assessment data (MoE 2019b) or PSCIS assessment data (MoE 2021) and associate with stream segment lines. When both FISS and PSCIS values were associated with a particular stream segment, or multiple FISS channel widths are available a mean of the average channel widths was used. To model channel width for 2nd order and above stream segments without associated FISS or PSCIS sites, first fwapg was used to estimate the drainage area upstream of the segment. Then, rasters from ClimateBC (Wang et al. 2012) were downloaded to a postgresql database, sampled for upstream watershed areas associated with each stream segment and a mean annual precipitation weighted by upstream watershed area was calculated. In early 2021, Bayesian statistical methods were developed to predict channel width in all provincial freshwater atlas stream segments where width measurements had not previously been measured in the field. The model was based on the relationship between watershed area and mean annual precipitation weighted by upstream watershed area (Thorley and Irvine 2021). In December of 2021, Thorley and Irvine (2021) methods were updated using on a power model derived by Finnegan et al. (2005) which relates stream discharge to watershed area and mean annual precipitation. Data (n = 24849) on watershed size, mean annual precipitation and measured channel width was extracted from the provincial freshwater atlas (FLNRORD 2021; GeoBC 2022), the BC Data Catalogue fisheries datasets (MoE 2020b, 2021) and Wang et al. (2012) utilizing bcfishpass (Norris [2020] 2021) and fwapg (Norris [2019] 2021). Details of this analysis and subsequent outputs can be reviewed [here](https://www.poissonconsulting.ca/f/859859031) (Thorley, Norris, and Irvine 2021).”_  Reference: BC Fish Passage [paper](https://a100.gov.bc.ca/pub/acat/documents/r62435/PEA_F23_F_3761_DCA_1697814406164_EFF767D4C4.pdf)

Channel width code - https://github.com/smnorris/fwapg/tree/main/extras/channel_width/sql

</details>

### Riparian Classification Methodology Summary

Layers were classified according to this FRPA [regulation guidelines](https://www.bclaws.gov.bc.ca/civix/document/id/complete/statreg/14_2004#division_d2e9829).

__Wetlands__
The FWA Wetland polygons served as the primary spatial layer for deriving area-based classifications. For the W2, W3, and W4 classifications, Biogeoclimatic Ecosystem Classification (BEC) attributes were required. To assign BEC zones and subzones to each wetland, a spatial intersection was performed between the BEC layer and the wetlands polygons. If a wetland intersected multiple BEC zones or subzones, the zone or subzone covering the largest portion of the wetland was selected.

__Lakes__
The FWA Lake polygons served as the primary spatial layer for deriving area-based classifications. For the L2, L3, and L4 classifications, Biogeoclimatic Ecosystem Classification (BEC) attributes were required. To assign BEC zones and subzones to each lake, a spatial intersection was performed between the BEC layer and the lake polygons. If a lake intersected multiple BEC zones or subzones, the zone or subzone covering the largest portion of the lake was selected.

__Streams__
The primary data sources used to derive stream classifications were the Fish Habitat Accessibility Model, FWA Rivers, Community Watersheds, the Mapped Floodplains in BC, and modeled channel width. The data format of the Fish Habitat Accessibility Model is a spatial linestring stream network for the province, identical to the FWA Stream Network but divided into smaller segments or reaches.

In order to assign stream riparian classes, the following pre-processing steps were performed:

* Each stream network is assigned a modeled channel width. See Data Sources for further description of the channel width model and its limitations. When channel width was unavailable, stream order was used as a proxy for channel width.
* Each stream network reach was assigned a value for community watershed presence: TRUE if the stream falls within a community watershed, and FALSE if it does not.
* Each stream network reach was assigned a value for active floodplain presence: TRUE if the stream falls within an active floodplain, and FALSE if it does not. All streams that fell within an active floodplain were manually inspected and assigned a value if the surrounding floodplain was >= 100m.
* Each stream is assigned a Natural Resource Area (ex. North, South, Coast)

For the majority of streams in the province, stream classification was assigned to each stream reach using the above added attributes. For streams that did not have modeled channel width, stream class was assigned based on stream order and natural resource area. 
<details>
  <summary> Expand for further details of how unclassified streams were handled.  </summary>
  <br>

Many streams have a NULL value for modeled channel width. This may occur either because the stream's catchment basin extends beyond BC or due to gaps in the modeled channel width dataset. Stream classification was divided by administrative areas, following advice from Erin Moore, who noted significant differences between coastal and interior regions.

__North and South Admin Areas__

Approach taken from 2024 Boundary TSA for North and South Admin Areas. 
Source: `file:///G:/!Workgrp/Analysts/!Project/TSA_Projects/active/KootenayLake_TSA13/TSR4/AnalysisReport/TSA13_TSR4_AnalysisReport/03-data-analysis.html#riparian-analysis`


```{r eval = TRUE, echo=FALSE}
library(knitr)
stream_data <- data.frame(
  `Stream Classification Assigned` = c("S1B (where not S1A)", "S2", "S3", "S4", "S5", "S6"),
  `Within Community Watershed OR Fish_Passage` = c(NA, "Yes", "Yes", "Yes", "No", "No"),
  `Stream Order` = c(">=4", "3", "2", "1", ">1", "1"),
  check.names=FALSE # to preserve column names with spaces
)

kable(stream_data, caption='Stream Classification Assignment within North and South Admin Area (when channel width model does not exist)')
```

__Coast Admin Area__

Approach taken from 2020 North Island TSA. Source:
https://www2.gov.bc.ca/assets/gov/farming-natural-resources-and-industry/forestry/stewardship/forest-analysis-inventory/tsr-annual-allowable-cut/48tsdp_2020.pdf

```{r eval = TRUE, echo=FALSE}
library(knitr)
stream_data <- data.frame(
  `Stream Classification Assigned` = c("S1B", "S2", "S3", "S4", "S5", "S6"),
  `Within Community Watershed OR Fish_Passage` = c("Yes", "Yes", "Yes", "Yes", "No", "No"),
  `Stream Order` = c(">=5", "4", "(3,2)", "1", ">1", "1"),
  check.names=FALSE # to preserve column names with spaces
)

kable(stream_data, caption='Stream Classification Assignment within Coastal Admin Area (when channel width model does not exist)')
```
</details>
<br>
__Rivers__
For a complete stream classification, the FWA Rivers polygon dataset is typically used in Timber Supply Analyses. It is important to include this dataset to ensure that river areas are accounted for in riparian zones. To classify each FWA River polygon, the most common stream classification (i.e., the mode) from the overlapping stream reaches was assigned to the polygon. As river polygons are known to be large rivers, it was assumed erroneous if any were classified as S6 (channel width <= 3m), S3 (channel width >= 1.5m and < 5m) or S4 (< 1.5m). Any FWA rivers classified as S6 were altered to S5 and any classified as S3 or S4 were changed to S2. 

### Buffers

Buffers were determined using the Riparian management area guidebook found [here](https://www2.gov.bc.ca/gov/content/industry/forestry/managing-our-forest-resources/silviculture/silvicultural-systems/silviculture-guidebooks/riparian-management-area-guidebook)  

__Restrictions in a riparian management zones__

> _"Under the Forest practices within the RMA, fisheries-sensitive zones, and marine-sensitive zones section. The following table summarizes the maximum overall levels of retention within the riparian management zones for each riparian class of stream, wetland and lake that are anticipated to result from consistent implementation of the best management practices recommended in this guidebook. "_

The following tables were copied from Riparian Management area Guidebook. The second last column (Maximum Overall Retention in the Riparian Management Zone (%). is The last column (I.e. Applied Riparian Buffer width (metres)) has been derived and is the result of (reserve zone + (management zone (m) * Maximum Overall Retention in the Riparian Management Zone (%))) / 2.

This analysis applies a riparian buffer with 100% retention to the entire Applied Riparian Buffer width, rather than following the recommended approach of retaining 100% in the Riparian Reserve Zone and applying a maximum overall retention in the Riparian Management Zone.

For example, an S2 stream would receive a single 40m buffer with 100% retention, instead of a two-tiered approach: a 30m buffer with 100% retention immediately around the stream, followed by an additional 20m buffer with 50% retention extending outward from the first buffer.

```{r echo = FALSE, results = 'asis'}
library(knitr)
## streams
riparian_class <- c("S1-A", "S1-B", "S2", "S3", "S4", "S5", "S6")
management_area <- c(100, 70, 50, 40, 30, 30, 20)
reserve_zone <- c(0, 50, 30, 20, 0, 0, 0)
management_zone <- c(100, 20, 20, 20, 30, 30, 20)
maximum_overall_retention <- c("50", "50", "50", "50", "25", "25", "5")
applied_riparian_buffers <- c(50, 60, 40, 30, 7.5, 7.5, 1)

stream_riparian_df <- data.frame(
  `Riparian Class` = riparian_class,
  `Riparian Reserve Zone (metres)` = reserve_zone,
  `Riparian Mgmt Zone (metres)` = management_zone,
 `Riparian Mgmt Area (metres)` = management_area,
  `Maximum Overall Retention in the Riparian Mgmt Zone (%)` = maximum_overall_retention,
  `Applied Riparian Buffer width (metres)` = applied_riparian_buffers,
  check.names = FALSE # to preserve column names with spaces
)
kable(stream_riparian_df, caption='Stream Riparian Buffer Guidelines')

## wetlands
riparian_class <- c("W1", "W2", "W3", "W4", "W5")
management_area <- c(50, 30, 30, 30, 50)
reserve_zone <- c(10, 10, 0, 0, 10)
management_zone <- c(40, 20, 30, 30, 40)
maximum_overall_retention <- c("25", "25", "25", "25", "25")
applied_riparian_buffers <- c(20, 15, 7.5, 7.5, 20)


wetland_riparian_df <- data.frame(
  `Riparian Class` = riparian_class,
  `Riparian Reserve Zone (metres)` = reserve_zone,
  `Riparian Mgmt Zone (metres)` = management_zone,
  `Riparian Mgmt Area (metres)` = management_area,
  `Maximum Overall Retention in the Riparian Mgmt Zone (%)` = maximum_overall_retention,
  `Applied Riparian Buffer width (metres)` = applied_riparian_buffers,
  check.names = FALSE # to preserve column names with spaces
)
kable(wetland_riparian_df, caption='Wetland Riparian Buffer Guidelines')

## lakes
riparian_class <- c("L1-A", "L1-B", "L2", "L3", "L4")
management_area <- c(0, 10, 30, 30, 30)
reserve_zone <- c(0, 10, 10, 0, 0)
management_zone <- c(0, 0, 20, 30, 30)
maximum_overall_retention <- c("25", "25", "25", "25", "25")
applied_riparian_buffers <- c(0, 10, 15, 7.5, 7.5)

lake_riparian_df <- data.frame(
  `Riparian Class` = riparian_class,
  `Riparian Reserve Zone (metres)` = reserve_zone,
  `Riparian Mgmt Zone (metres)` = management_zone,
  `Riparian Mgmt Area (metres)` = management_area,
  `Maximum Overall Retention in the Riparian Mgmt Zone (%)` = maximum_overall_retention,
  `Applied Riparian Buffer width (metres)` = applied_riparian_buffers,
  check.names = FALSE # to preserve column names with spaces
)
kable(lake_riparian_df, caption='Lake Riparian Riparian Buffer Guidelines')


```

The buffered areas were converted into a numeric raster, where each cell represents the proportion of the cell covered by the buffer.



### R Scripts
The methodology used an integrated approach using R and PostGIS. R scripts were used to construct and execute SQL queries on the PostGIS database. The heavy spatial operations were performed directly within the PostGIS database, leveraging its robust spatial functions and indexes. 

The following R scripts were used to import all needed data sources, post process imported datasets, perform riparian classification and buffer the riparian classifications.

```
src\analysis\1.riparian-import-layers.R
src\analysis\2.riparian-fwapg-pre-processing.R
src\analysis\3.riparian-classification.R
```


### Known Limitations
<details>
  <summary> Expand for known limitations of provincial modeled lake, wetland and stream classification </summary>
  <br>
  
__Usage Limitations__

__Stream__ - Riparian stream classification can only be assigned if the channels have been measured and sampled for fish presence or absence, which requires field work. As such, no provincial riparian classification database exists. This stream classification model is __not appropriate for operational use__. It is acceptable for generating broad landscape level summary stats for something like AAC Impacts and TSR.

__General Limitations__

Lisa Nordin, Aquatic Resources Stewardship Evaluation Office, was contacted to get clarification on regulation. During the email exchange, she identified common misclassifications and other inherent problems with using modeled data and the FWA.


> _"The FWA is unreliable when it comes to water features, especially streams. They are vastly over-represented in the interior and under-represented on the coast."_

> _"I have strong concerns over what might be considered non fish (S5 or S6). Under current regulation it is acceptable to say anything over 20% gradient is non-fish (even though most biologists will tell you certain species can move through those gradients with ease, but except for rare circumstances they aren't likely to reside there). However, it would be incorrect to say everything above a steep gradient or identified barrier is also non-fish if it is less than 20% unless it has been sampled, or field confirmed that all upstream reaches are dry or frozen to the bottom at any time of year. I am not sure what the model is telling you but please do not make that mistake -  when you are mapping it is critical that all reaches less than 20% be identified as a fish stream unless classed otherwise using ground based methods. This is in our current regulation and goes back to the forest practices code days so it isn't new. I can't stress how important this is as we have been seeing misclassifications and subsequent harvesting of reserves. I know this because I have been an expert witness in legal cases where a licensee saw a waterfall and called everything upstream non-fish bearing and harvested it to the edge without sampling. The Forest Practices Board is also tuned into this and runs audits to make sure there is appropriate sampling above barriers to prove a non-fish status.   To try to correct for this, we are currently in the last stages of revising and posting the fish stream identification guidebook and in the meantime posted this 3-page Extension Note [here](https://www2.gov.bc.ca/assets/gov/farming-natural-resources-and-industry/forestry/frep/extension-notes/frep-fish-stream-identifcation_final.pdf)."_

__Wetland__ - Riparian wetland classification relies on the FWA wetlands layer. 

Identified problems with using the FWA wetlands for classify riparian zones for wetlands by Lisa Nordin:

> _"Non classified wetlands (NCWs) are not included management areas. However, it should be noted that in many (many!) cases, NCWs were incorrectly identified as being separate when in fact they are not - sometimes because of a treed section between them and the next wetland that makes a GIS user believe it is not a wetland type, or incorrect spatial vegetation layers. This was occurring so frequently, webinars for practioners were ran on wetland identification and delineation, [link](https://www.fpbc.ca/professional-development/continuing-professional-development/offerings/webinar-recordings/) . The video describes the value of ground verification and site indicators, but there is also some good info on spatial sources of information."_

Lisa is correct to have these concerns about the classification of S5 and S6 streams and non-classified wetlands from an operational perspective. Hence the usage limitations for this modeled dataset.

__Analysis Methodology Limitations__

__FWA River Polygons__ - A review of several TSAs was conducted to understand how river polygons are classified. The guidelines specify that a stream is classified as S1A if it averages, over a one-kilometer length, either a stream width or an active floodplain width of 100 meters or more. A stream is classified as S1B if its width exceeds 20 meters but it does not meet the criteria for S1A. Since TSAs often lacked access to channel width data, they typically classified river polygons S1A or S1B by manually measuring features from the river polygon or satellite imagery. As this analysis has access to modelled channel width, that was leveraged to classify the river polygons.

The following approach was used:

1. Each stream reach is classified individually.
2. The stream reaches within a river polygon are identified
3. The river polygon is assigned the classification of the most common riparian class (I.e., mode) of the stream reaches that it overlaps.

As FWA Rivers are known to be large rivers, it was assumed erroneous if any were classified as S6 (channel width <= 3m), S3 (channel width >= 1.5m and < 5m) or S4 (< 1.5m). Any FWA rivers classified as S6 were altered to S5 and any classified as S3 or S4 were changed to S2. 

Many river polygons were not classified because they lack a stream reach within them. This commonly occurs in braided rivers, where a smaller river polygon is attached to a larger polygon but lacks any stream network linework. In these instances, the FWA polygon was assigned the lowest classification among all intersecting FWA polygons.

This approach is considered a limitation as it is new and differs from TSA analysis and has not been checked against field data.

</details>

### References and resources
<details>
  <summary> Expand for an overview of references and resources used to guide classification </summary>
  <br>

__Regulation__
The riparian classification requirements were primarily guided by the Forest and Range Practices Act, Forest Planning and Practices Regulation (FPPR) [link](https://www.bclaws.gov.bc.ca/civix/document/id/complete/statreg/14_2004#division_d2e9829). Specific subsection: Division 3 - Riparian Areas.

__Timber Supply Area Examples__

Rhian's example
file:///G:/!Workgrp/Analysts/!Project/TSA_Projects/active/KootenayLake_TSA13/TSR4/AnalysisReport/TSA13_TSR4_AnalysisReport/03-data-analysis.html#riparian-analysis

__Wetland additional notes__: When classifying W5 wetlands, the regulation states:

> _"(a) the area contains:  
  (i) two or more W1 wetlands located within 100 m of each other,  
  (ii) a W1 wetland and one or more non-W1 wetlands, all of which are within 80 m of each other, or  
  (iii) two or more non-W1 wetlands located within 60 m of each other, and  
(b) the combined size of the wetlands, excluding the upland areas, is 5 ha or larger."_


It was unclear whether "non-W1 wetlands" included non-classified wetlands and unclear whether bullet (i) should be read as having an "or" at the end of the line. Clarification was given by Lisa Nordin:

> _"The FPPR definitions carried over from the Code and are based on those in the FPC Riparian Management Area Guidebook (which is no longer referenced in our regulation). The W5 class is a wetland complex meaning there are 2 or more individual wetlands with overlapping RMAs. The FPPR s.48(2) is just outlining the distance when that would occur. So, a W1 has an RMA of 50. To have two of them overlap, you would need a distance of less than 100m between them.  W2, W3, and W4 wetlands have RMAs of 30. So if you had one W1 and any one of the others, then 80m is your max overlap distance. Two of the smaller wetlands together would be 60m, but the polygon (including the distance between them) has to be more than 5 ha for it to be a W5, hence the "and" after iii. 
To answer your second questions, unfortunately NCWs are not included in the W5 classification because they do not have an RMA associated with them."_

As such, W5's were assigned using the following workflow:

* buffer's were created on polygons classified as W1 (50m), W2 (30m), W3 (30m) or W4 (30m)  
* all buffers that overlapped with one or more other buffers were identified and grouped together  
* the sum of the wetland area of all identified overlapping groups was calculated  
* if the area was >= 5 ha, W5 was assigned  

</details>



#### QA/QC
<details>
  <summary> Expand for QA/QC </summary>
  <br>

Compare the modeled stream width with FREP data
```{r eval = FALSE}
query <- "SELECT
min(frep.distance_to_line),
avg(frep.distance_to_line),
stddev(frep.distance_to_line),
max(frep.distance_to_line)
FROM
thlb_proxy.modelled_habitat_potential stream
JOIN
thlb_proxy.june13_riparian_data_for_faib frep
ON
frep.modelled_habitat_potential_fid = stream.fid
WHERE stream.channel_width is not null
and
stream.riparian_class != frep.stream_class_in_field   -- count = 904, min = 0.004, avg = 97, stddev = 169, max = 1151
-- stream.riparian_class = frep.stream_class_in_field -- count = 880, min = 0.009, avg = 37, stddev = 83,  max = 716"


query <- "SELECT
	stream.fid,
	frep.objectid,
	stream.slope,
	stream.channel_width,
	frep.channel_width as width_in_field,
	frep.channel_gradient_pct,
	stream.riparian_class,
	frep.stream_class_in_field
FROM
	thlb_proxy.modelled_habitat_potential stream
JOIN
		thlb_proxy.june13_riparian_data_for_faib frep
ON
frep.modelled_habitat_potential_fid = stream.fid
WHERE 
	stream.channel_width is not null
AND
stream.riparian_class != frep.stream_class_in_field -- count = 904
--stream.riparian_class = frep.stream_class_in_field -- count = 880"
```

</details>

#### QA/QC
<details>
  <summary> Expand for QA/QC review </summary>
  <br>
  
__Wetland QA/QC__
Visually compared results with Kootenay Lakes TSA results. Almost identical except for 2 cases:

+ In TSR, two sets of nearby wetlands were classified as W5, but above analysis calculated W5. I think its mis-classified as the summed area was < 5. (Eg., waterbody_poly_id in (705016913, 705016826) and (705016771, 705016732))


__Lake QA/QC__
Visually compared the above output with Rhian's layer for Kootenay Lakes TSA (TSA13) - found to be identical.
Rhian's layer: W:/FOR/VIC/HTS/ANA/Workarea/ardavies/ts_units/2019/tsa13/Tasks/005_riparian/data/riparian.gdb|layername=lake_classes

__Stream QA/QC__
Visually compared the above output with Rhian's layer for Kootenay Lakes TSA (TSA13) - there were many differences as Rhian used a stream order based approach and this analysis used modeled channel width. 


### Rasterize the buffers
<details>
  <summary> Convert the rasters into a gr_skey lookup table </summary>
  <br>
